County Figures

### 2023-06-26 reports/county_figures.Rmd
improve and add county-level figures

library(tidyverse)
library(targets)
library(here)
library(latex2exp)
library(kableExtra)

@@

waste <- tar_read(waste, store =targets_dir)
covidestim <- tar_read(covidestim_biweekly_county,
                        store =targets_dir)

@@

wjoined <- county %>%
  # set Nantucket, Duke fips to Nantucket since we have wastewater data
  # for Nantucket
  mutate(fips = ifelse(grepl("25019", fips), "25019", fips)) %>%
  inner_join(waste) %>%
  left_join(covidestim, relationship = "many-to-many") %>%
  group_by(biweek) %>%
  mutate(mindate = min(date),
         maxdate = max(date)) %>%
  ungroup()

@@

w_data%>%
  mutate(name = gsub("County, MA", "", name)) %>%
  left_join(dates,relationship = "many-to-many" ) %>%
  filter(version %in% c("v1","v3")) %>%
  filter(biweek>=6) %>%
  group_by(biweek) %>%
  mutate(mindate = min(date),
         maxdate=max(date)) %>%
  ungroup() %>% 
 #   filter(fips == county_fips  & date >= begin_date & date <= end_date) %>%
    ggplot() +
      geom_ribbon(aes(x = date, 
                 ymin = exp_cases_lb_std,
                 ymax = exp_cases_ub_std,
                 fill = vlabel),
                 alpha = .5) +
      geom_line(data = waste_df,
                aes(x = date, y =mean_conc_std, 
                    color = 'Wastewater Effective Concentrations'),
               # color = "#DB4048",
                size = 1.1) +
      geom_linerange(aes(xmin = mindate, 
                         xmax = maxdate, 
                         y = infections_std,
                         color = 'Covidestim Infection Estimates')) + 
      geom_point(data = waste_df,
                aes(x = date, y =mean_conc_std ),
                color = "#DB4048",
                alpha = .5,
                size = 1.2) +
      facet_grid(name~vlabel, 
                 labeller=labeller(vlabel =as_labeller(
                   TeX, default=label_parsed))) +
      scale_fill_manual(values = pal, labels = c('','','',''), 
                        name = "Probabilistic Bias Intervals") +
      theme_bw() +
     theme_c() +
      labs(y = "Value (Standardized by County)",
           title = "",
           x= "") +
      scale_x_date(date_labels = "%b %Y") +
    scale_color_manual(
      name = '', 
      values = c(
        'Covidestim Infection Estimates' = 'darkblue',
        'Wastewater Effective Concentrations' = '#DB4048')) +
    guides(color = guide_legend(override.aes = list(size = 12,
                                                    linewidth=3)),
           fill = guide_legend(override.aes =list(size = 6)))

@@

@@

Although the spline-smoothing improved the correlations, it did not improve the proportion containing the Covidestim median.

>>>>

library(tidyverse)
library(targets)
library(here)
library(latex2exp)
library(kableExtra)
library(scales)
library(cowplot)
```

@@

waste <- tar_read(waste, store =targets_dir)
county_names <- read_tsv(here("data/demographic/county_fips.tsv")) %>%
  rename_with(.cols = everything(), .fn =tolower) %>%
  bind_rows()

covidestim <- tar_read(covidestim_biweekly_county,
                        store =targets_dir) %>%
  select(-c(date)) %>% 
  distinct()

@@

wjoined <- county %>%
  # set Nantucket, Duke fips to Nantucket since we have wastewater data
  # for Nantucket
  mutate(fips = ifelse(grepl("25019", fips), "25019", fips)) %>%
  inner_join(waste) %>%
  left_join(covidestim, relationship = "many-to-many") %>%
  left_join(dates,  relationship = "many-to-many") %>%
  group_by(biweek) %>%
  mutate(mindate = min(date),
         maxdate = max(date)) %>%
  ungroup()

@@

w_data <- w_data %>%
  group_by(name) %>%
  mutate(adjust = 1/(max(mean_conc))*(exp_cases_median[which.max(mean_conc)])) %>%
  ungroup()

waste_df <- w_data %>%
  select(biweek, mean_conc, name, adjust, fips) %>%
  distinct() %>%
  left_join(dates,relationship = "many-to-many" ) %>%
  group_by(name, biweek) %>%
  # center date within 2-week interval
  summarize(date = min(date) + days(7),
            mean_conc =unique(mean_conc),
            adjust=unique(adjust),
            fips = unique(fips)) %>%
  mutate(name = gsub("County, MA", "", name))

pltlist <- w_data %>%
  mutate(name = gsub("County, MA", "", name)) %>%
  left_join(dates,relationship = "many-to-many" ) %>%
  # filter(version %in% c("v1","v3")) %>%
  filter(version=="v1") %>%
  filter(biweek>=6) %>%
  group_by(biweek) %>%
  mutate(mindate = min(date),
         maxdate=max(date)) %>%
  ungroup() %>% 
  group_by(name) %>%
  group_split() %>%
  map(~ {

    county_fips <- .x$fips %>% unique()
    adj <- .x$adjust %>% unique()

    .x %>% 
 #   filter(fips == county_fips  & date >= begin_date & date <= end_date) %>%
    ggplot() +
      geom_ribbon(aes(x = date, 
                 ymin = exp_cases_lb,
                 ymax = exp_cases_ub,
                 fill = vlabel),
                 alpha = .5) +
      geom_line(data = waste_df[which(waste_df$fips == county_fips),],
                aes(x = date, y =mean_conc*adj, 
                    color = 'Wastewater Effective Concentrations'),
               # color = "#DB4048",
                size = 1.1) +
      geom_linerange(aes(xmin = mindate, 
                         xmax = maxdate, 
                         y = infections,
                         color = 'Covidestim Infection Estimates')) + 
      geom_point(data =  waste_df[which(waste_df$fips == county_fips),],
                aes(x = date, y =mean_conc*adj ),
                color = "#DB4048",
                alpha = .5,
                size = 1.2) +
      scale_y_continuous(sec.axis=sec_axis(~.x/adj,
                                           name = 'Mean Wastewater Effective Concentration',
                                           labels=comma),
                         labels=comma
                    ) +
      # facet_grid(name~vlabel, 
      #            labeller=labeller(vlabel =as_labeller(
      #              TeX, default=label_parsed))) +
      facet_wrap(~name, ncol=3) + 
      scale_fill_manual(values = pal, labels = c('','','',''), 
                        name = "Probabilistic Bias Intervals") +
     theme_c(legend.position="none") +
      labs(y = "Value (Standardized by County)",
           title = "",
           x= "") +
      scale_x_date(date_labels = "%b %Y") +
    scale_color_manual(
      name = '', 
      values = c(
        'Covidestim Infection Estimates' = 'darkblue',
        'Wastewater Effective Concentrations' = '#DB4048')) +
    guides(color = guide_legend(override.aes = list(size = 12,
                                                    linewidth=3)),
           fill = guide_legend(override.aes =list(size = 6))) 
    })

legend <- cowplot::get_legend(
  pltlist[[1]] + theme(legend.position="top")
)

grid <- cowplot::plot_grid(plotlist=pltlist, ncol =3)

cowplot::plot_grid(legend, grid, nrow=2, rel_heights=c(.05,.95))

```{r, eval=FALSE,include=FALSE}

 w_data_std <- w_data %>%
  group_by(name, version) %>%
  mutate(exp_cases_lb = exp_cases_lb /max(exp_cases_median),
         exp_cases_ub = exp_cases_ub /max(exp_cases_median),
         exp_cases_median = exp_cases_median / max(exp_cases_median),
         mean_conc=mean_conc/max(mean_conc),
         infections=infections/max(infections)) %>%
  mutate(name = gsub("County, MA", "", name)) %>%
  left_join(dates,relationship = "many-to-many" ) %>%
  filter(version %in% c("v1","v3")) %>%
  filter(biweek>=6) %>%
  group_by(biweek) %>%
  mutate(mindate = min(date),
         maxdate=max(date)) %>%
  ungroup()

waste_df <- w_data_std %>%
  select(biweek, mean_conc, name, fips) %>%
  distinct() %>%
  left_join(dates,relationship = "many-to-many" ) %>%
  group_by(name, biweek) %>%
  # center date within 2-week interval
  summarize(date = min(date) + days(7),
            mean_conc =unique(mean_conc),
            fips = unique(fips)) %>%
  mutate(name = gsub("County, MA", "", name))

w_data_std %>% 
 #   filter(fips == county_fips  & date >= begin_date & date <= end_date) %>%
    ggplot() +
      geom_ribbon(aes(x = date, 
                 ymin = exp_cases_lb,
                 ymax = exp_cases_ub,
                 fill = vlabel),
                 alpha = .5) +
      geom_line(data = waste_df,
                aes(x = date, y =mean_conc, 
                    color = 'Wastewater Effective Concentrations'),
               # color = "#DB4048",
                size = 1.1) +
      geom_linerange(aes(xmin = mindate, 
                         xmax = maxdate, 
                         y = infections,
                         color = 'Covidestim Infection Estimates')) + 
      geom_point(data =  waste_df,
                aes(x = date, y =mean_conc),
                color = "#DB4048",
                alpha = .5,
                size = 1.2) +
   #   scale_y_continuous(sec.axis=sec_axis(~.x/adj)) +
      facet_grid(name~vlabel, 
                 labeller=labeller(vlabel =as_labeller(
                   TeX, default=label_parsed))) +
      scale_fill_manual(values = pal, labels = c('','','',''), 
                        name = "Probabilistic Bias Intervals") +
      theme_bw() +
     theme_c() +
      labs(y = "Value (Standardized by County)",
           title = "",
           x= "") +
      scale_x_date(date_labels = "%b %Y") +
    scale_color_manual(
      name = '', 
      values = c(
        'Covidestim Infection Estimates' = 'darkblue',
        'Wastewater Effective Concentrations' = '#DB4048')) +
    guides(color = guide_legend(override.aes = list(size = 12,
                                                    linewidth=3)),
           fill = guide_legend(override.aes =list(size = 6)))

@@

# Simulation Intervals Over Time

```{r, fig.width=11, fig.height=8}

county_dat <- county %>%
  # set Nantucket, Duke fips to Nantucket since we have wastewater data
  # for Nantucket
  left_join(dates,  relationship = "many-to-many") %>%
  group_by(biweek) %>%
  mutate(mindate = min(date),
         maxdate = max(date)) %>%
  ungroup() %>%
  left_join(county_names, by=c('fips'='fips')) %>%
  mutate(name = ifelse(fips =="25007,25019", "Nantucket and Dukes", name))

county_dat %>%
  filter(version=="v1")  %>%
   ggplot() +
  geom_ribbon(aes(x = date, 
             ymin = exp_cases_lb,
             ymax = exp_cases_ub),
             alpha = .5,
             show.legend=FALSE,
             fill="#596281") +
   geom_linerange(aes(xmin = mindate,
                      xmax=maxdate,
                      y = positive,
                      color = 'obs'),
                  size=.5) +
  facet_wrap(~name, scales="free_y", ncol=3) +
  scale_y_continuous(labels = comma) +
  scale_x_date(date_breaks = "3 months", 
               date_labels = "%b %Y") +
  theme_c(
          legend.position = "top",
          ) +
  scale_fill_manual(values = pal)  +
  labs(x="",
       y = "Estimated Infections",
       title = "Simulation Intervals Over Time",
       subtitle = "Counties in Massachusetts") +
  scale_color_manual(values = c('obs' = 'red'), labels = 'Positive Tests',
                     name = '') +
  guides(color = guide_legend(override.aes = list(linewidth =2)))

## Ratio of Estimated Infections to Observed

```{r, fig.width=12, fig.height=9}

county_dat %>%
  filter(version=="v1")  %>%
  #filter(grepl("Nan",name)) %>%
   ggplot() +
  geom_ribbon(aes(x = date, 
             ymin = exp_cases_lb/positive,
             ymax = exp_cases_ub/positive),
             alpha = .5,
             show.legend=FALSE,
             fill="#596281") +
   # geom_linerange(aes(xmin = mindate,
   #                    xmax=maxdate,
   #                    y = positive,
   #                    color = 'obs')) +
  facet_wrap(~name,ncol=3) +
 # facet_grid(name~vlabel, 
 #                 labeller=labeller(vlabel =as_labeller(
 #                   TeX, default=label_parsed))) +
  scale_y_continuous(labels = comma) +
  scale_x_date(date_breaks = "3 months", 
               date_labels = "%b %Y") +
  theme_c(
          legend.position = "top") +
  scale_fill_manual(values = pal)  +
  labs(x="",
       y = "Estimated Infections",
       title = "Ratio of Estimated to Observed Over Time",
       subtitle = "Counties in Massachusetts") +
  scale_color_manual(values = c('obs' = 'red'), labels = 'Positive Tests',
                     name = '') +
  guides(color = guide_legend(override.aes = list(linewidth =2)))

# Trends in Total Tests and Positive Tests

Informs our understanding of the trends in the ratio of estimated to observed infections.

## Positive Tests and Test Positivity

```{r, fig.height=12, fig.width=12}

############################################
# POSITIVE TESTS AND TEST POSITIVITY
############################################
pltlist <- county_dat %>%filter(version=="v1") %>%
    mutate(testpos = positive/total) %>%
  group_by(across(-date)) %>%
  summarize(date=min(date) + days(7)) %>%
  group_by(fips) %>%
  mutate(adj = (1/max(testpos)) * max(positive)) %>%
  group_split() %>%
  map(~ {
    adj <-unique(.x$adj)
    .x %>%
      ggplot(aes(x=date))+
    geom_line(aes(y=positive, color = 'Positive Tests')) +
      geom_point(aes(y=positive, color = 'Positive Tests'),
              alpha=.5,
              size=.8) +
    geom_line(aes(y=testpos*adj, color='Test Positivity'),
              linetype=2) +
      geom_point(aes(y=testpos*adj, color='Test Positivity'),
              alpha=.5,
              size=.8) +
    #scale_linetype_discrete(breaks = c(1,1,2)) +
    scale_color_manual(values=c(
                                "Positive Tests"= "#279143",
                                "Test Positivity"= '#E38E4F')) +
    guides(color = guide_legend(override.aes = list(linetype = c(1,2),
                                                    linewidth=c(2,2)))) +
    theme_c() +
    theme(axis.title = element_text(size=10),
          axis.text =element_text(size=9),
          legend.position="none") +
    scale_y_continuous(labels=comma, sec.axis = sec_axis(~./adj , name = 'Test Positivity')) +
    labs(color ='',
         y="Positive Tests",
         x= "Date") +
      facet_wrap(~name)

  })

legend <- get_legend(
  pltlist[[1]] +
    theme(legend.position = "top",
          legend.text=element_text(size=16))
)

title_gg <- ggplot() + 
  labs(title = "Comparing Trends in Positive Tests and Test Positivity") + 
  theme(plot.title=element_text(face="bold",
                                hjust = .5, 
                                size = 18,
                                margin =margin(5,0,2,0)))

plt <- cowplot::plot_grid(plotlist=pltlist, ncol=3)

plot_grid(title_gg, legend, plt, rel_heights=c(.03, .03,.94), ncol=1)

## Total Tests and Positive Tests

```{r, fig.height=12, fig.width=12}

#####################################
# TOTAL TESTS AND POSITIVE TESTS
#####################################
pltlist <- county_dat %>%filter(version=="v1") %>%
    mutate(testpos = positive/total) %>%
  group_by(across(-date)) %>%
  summarize(date=min(date) + days(7)) %>%
  group_by(fips) %>%
  mutate(adj = (1/max(total)) * max(positive)) %>%
  group_split() %>%
  map(~ {
    adj <-unique(.x$adj)
    .x %>%
      ggplot(aes(x=date))+
    geom_line(aes(y=positive, color = 'Positive Tests')) +
      geom_point(aes(y=positive, color = 'Positive Tests'),
              alpha=.5,
              size=.8) +
    geom_line(aes(y=total*adj, color='Total Tests'),
              linetype=2) +
      geom_point(aes(y=total*adj, color='Total Tests'),
              alpha=.5,
              size=.8) +
    #scale_linetype_discrete(breaks = c(1,1,2)) +
    scale_color_manual(values=c(
                                "Positive Tests"= "#279143",
                                "Total Tests"= '#E38E4F')) +
    guides(color = guide_legend(override.aes = list(linetype = c(1,2)))) +
    theme_c() +
    theme(axis.title = element_text(size=11),
          axis.text =element_text(size=8),
          legend.position="none") +
    scale_y_continuous(labels=label_scientific(), 
                       n.breaks=4,
                       sec.axis = sec_axis(~./adj ,
                                           name = 'Total Tests',
                                           labels=label_scientific())) +
    labs(color ='',
         y="Positive Tests",
         x= "Date") +
      facet_wrap(~name)
})

legend <- cowplot::get_legend(
  pltlist[[1]] +
  theme(legend.position="top") +
    guides(color = guide_legend(keywidth=2,
                                override.aes = list(linewidth=c(2,2),
                                                    linetype =c(1,2))))
)

title_gg <- ggplot() + 
  labs(title = "Comparing Trends in Total Tests and Positive Tests") + 
  theme(plot.title=element_text(face="bold",
                                hjust = .5, 
                                size = 18,
                                margin =margin(5,0,2,0)))

plt <- cowplot::plot_grid(plotlist=pltlist, ncol=3)

cowplot::plot_grid(title_gg, legend, plt, rel_heights =c(.03, .03,.94),ncol=1)

## Percent Change in Total Tests and Positive Tests

###################################################
# PERCENT CHANGE: TOTAL TESTS AND POSITIVE TESTS
###################################################

county_dat %>%
  filter(version=="v1") %>%
  filter(!grepl("Nantucket",name)) %>%
    mutate(testpos = positive/total) %>%
  group_by(across(-date)) %>%
  summarize(date=min(date) + days(7)) %>%
  group_by(fips) %>%
  arrange(biweek) %>%
  mutate(pct_change_positive = (positive - lag(positive, 1))/lag(positive, 1),
         pct_change_total = (total - lag(total, 1))/lag(total, 1)) %>%

   ggplot(aes(x=date))+
    geom_line(aes(y=pct_change_positive, color = 'Positive Tests')) +
      geom_point(aes(y=pct_change_positive, color = 'Positive Tests'),
              alpha=.5,
              size=.8) +
    geom_line(aes(y=pct_change_total, color='Total Tests'),
              linetype=2) +
      geom_point(aes(y=pct_change_total, color='Total Tests'),
              alpha=.5,
              size=.8) +
    #scale_linetype_discrete(breaks = c(1,1,2)) +
    scale_color_manual(values=c(
                                "Positive Tests"= "#279143",
                                "Total Tests"= '#E38E4F')) +
    guides(color = guide_legend(override.aes = list(linetype = c(1,2)))) +
    theme_c() +
    theme(axis.title = element_text(size=15),
          axis.text.x=element_text(size=9),
          legend.position="top") +
   # scale_y_continuous(labels=comma, sec.axis = sec_axis(~./adj , name = 'Test Positivity')) +
    labs(color ='',
         y="Percent Change",
         x= "Date",
         title = "Percent Change in Positive Tests and Total Tests") +
      facet_wrap(~name, ncol=3, scales="free_x") +
  scale_y_continuous(limits = c(-1,5.5),
                     labels=percent) +
  scale_x_date(date_breaks="3 months", date_labels="%b %Y")

@@

Although the spline-smoothing improved the correlations with wastewater concentrations and the correlations with Covidestim estimates, it did not improve the proportion containing the Covidestim median.
